Hybrid Threads for the Parrot Virtual Machine

### 2012-06-05 Paper.tex
Finish of the finish

@@

The initial goal of an almost lock free implementation has not been met fully. The garbage collector has to take a lock in most of it's functions. Though these locks are used only when more than one interpreter thread is running, they still represent an overhead which could be avoided. A way to accomplish this would be to pre-allocate proxy objects instead of allocating them on demand. The allocation could be done as part of a thread's outer runloop. This would eliminate the need for the locks in the garbage collector. The number of proxy objects to pre-allocate would be an important knob for optimization, since too few objects would lead to starvation quickly when many tasks are scheduled.

@@

@@

>>>>

Practice will show how this threading model will fare with real world workloads. It very much depends on the ratio of read to write accesses to shared data.

@@

The initial goal of an almost lock free implementation has not been met fully. The garbage collector has to take a lock in most of its functions. Though these locks are used only when more than one interpreter thread is running, they still represent an overhead which could be avoided. A way to accomplish this would be to pre-allocate proxy objects instead of allocating them on demand. The allocation could be done as part of a thread's outer runloop. This would eliminate the need for the locks in the garbage collector. The number of proxy objects to pre-allocate would be an important knob for optimization, since too few objects would lead to starvation quickly when many tasks are scheduled.

@@

The maximum number of threads is determined by the MAX\_THREADS constant in \textit{include/parrot/thread.h}. Instead the value should be determined at runtime probably depending on the available hardware resources, demand and system load.

@@

Currently the scheduler picks the interpreter with the smallest number of tasks in the task queue for scheduling a new task. This can lead to the situation where threads sit idle while tasks have to wait in other thread's queues. Some tasks get scheduled on specific threads because they need write access to the thread's PMCs. To allow migration of waiting tasks to idle threads there needs to be a method to distinguish such tasks from tasks which are there simply because the thread's queue was the shortest at the time the task got scheduled. The more difficult part of implementing task migration is that proxies are allocated on the target interpreter at the time of scheduling. Since these proxies belong to the interpreter they are created on, new proxies would have to be created on the interpreter the task is being migrated to. It is yet unclear if this overhead can be offset by better utilization of idle threads.

\section{Experiences}

I picked a very advanced topic for my Bachelor's thesis. While I had to invest many more hours than strictly necessary to get a Bachelor's degree, the work was also very interesting, instructive and hopefully valuable to the Parrot project and the free software community. Considering that I had no prior experience with Parrot which is a highly complex piece of software, it is not surprising that I spent most of the time discovering how it works and how its many features interact. In addition, concurrency usually brings with it difficult to diagnose problems further complicated by Parrot's cross platform nature.

It is therefore no exaggeration to state that without the help of the Parrot community I could not have come nearly as far as I did in the time I had. They are very welcoming and were supportive and grateful from the first day, boosting my motivation. I can only recommend to other students to approach the free software communities when looking for thesis topics.
